---
title: '*cool project name here*'
output:
  html_document:
    df_print: paged
authors: todos
---

----
libraries here 
```{r}
library(readxl)
library(classdata)
library(tidyverse)
library(stringr)
library(ggplot2)
library(plotly)
```

----
import data here
```{r}
fight_songs = read.csv('fight-songs.csv')
ipeds <- read_excel('IPEDS_data.xlsx', sheet='Data')
lookup_table = read.csv('lookup_table.csv')
```
----

cleaning data (VERSION 1)

1) loop through all values ipeds and fight_songs

2) add if similar name, otherwise print which names did not work

```{r}
# cleaned_ipeds <- data.frame()
# i <- 1
# for (name in fight_songs$school){
#   if (!is.na(pmatch(name, ipeds$Name, duplicates.ok = FALSE))){
#     picked_row <- (pmatch(name, ipeds$Name))
#     foo <- paste("works: ", name)
#     foo <- paste(foo, ipeds$Name[picked_row])
#     print(foo)
#    
#     cleaned_ipeds <- rbind(cleaned_ipeds, ipeds[picked_row,] )
#   }
#   else{
#     foo <- paste("doesn't work: ", name)
#     # print(foo)
#   }
# }
# 
# for(row in 1:nrow(lookup_table)){
#   location <- lookup_table$Column[row]
#   i <- i + 1
#   cleaned_ipeds <- rbind(cleaned_ipeds, ipeds[location,] )
# }

```
----
cleaning the data (VERSION 2!)
added a new column for the IDs in ipeds. allows for better, through cleaning. Some words weren't used correctly in pmatch.
```{r}
fight_songs
ipeds
cleaned_ipeds <- data.frame()

for (id in fight_songs$"ID.number"){
  if (!is.na(match(id, ipeds$"ID number"))){
    picked_row <- (match(id, ipeds$"ID number"))
    foo <- paste("works: ", id)
    print(foo)
    cleaned_ipeds <- rbind(cleaned_ipeds, ipeds[picked_row,] )
  }
  else{
    print("ERROR")
  }
}

names(fight_songs)[28] <- "ID number"
fight_songs$"ID number" = as.double(fight_songs$"ID number")
```

----
joining the data
```{r}
cleaned_ipeds$year <- NULL
join_data <- left_join(cleaned_ipeds, fight_songs, by="ID number")
```

----
jessies part
```{r}
join_data2 <- left_join(fight_songs, cleaned_ipeds)

colnames(join_data2)

speed_plot <- join_data2 %>%
  mutate('conference' = as.factor(join_data2$conference)) %>%
  ggplot(aes(x = sec_duration, y = bpm, color = join_data2$conference, text = paste0("<b>Song Name: </b>", join_data2$song_name, "<br>", "<b>School: <b>", join_data2$school))) +
  geom_point() +
  geom_hline(yintercept = median(join_data2$bpm)) +
  geom_vline(xintercept = median(join_data2$sec_duration)) +
  xlab("Song Length (Seconds)") +
  ylab("Song Speed (bpm)") +
  scale_colour_brewer(palette = "Set1") +
  theme(panel.grid = element_blank())

speed_plotly <- ggplotly(speed_plot, tooltip = "text")

speed_plotly

join_data2 <- join_data2 %>%
  mutate(speed_cat = if_else(join_data2$bpm < 140 & join_data2$sec_duration < 67, 'Short & Slow', if_else(join_data2$bpm < 140 & join_data2$sec_duration > 67, 'Long & Slow', if_else(join_data2$bpm > 140 & join_data2$sec_duration < 67, 'Short & Fast', 'Long & Fast')))) 

map <- map_data('state')

join_data2$year <- as.numeric(as.character(join_data2$year))

filter1900 <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1900)

map %>%
  ggplot(aes(x = long, y = lat)) + 
  geom_path(aes(group = group)) + 
  geom_point(data = filter1900, 
             aes(x = filter1900$'Longitude location of institution', y = filter1900$'Latitude location of institution', color = filter1900$student_writer)) +
  geom_label(data = filter1900, aes(filter1900$'Longitude location of institution', filter1900$'Latitude location of institution', label = year, color = filter1900$student_writer), hjust = 0, nudge_x = 0.3, check_overlap = TRUE) +
  scale_colour_brewer(palette = "Set1") +
  theme(panel.grid = element_blank()) +
  coord_map() +
  theme(legend.position = 'none', axis.text.x = element_blank(),
  axis.text.y = element_blank(), axis.ticks = element_blank()) +
  ylab("") +
  xlab("")

filter1905 <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1910 & join_data2$year > 1900)

filter1905T <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1910)

filter1905T$student_writer <- factor(filter1905T$student_writer, levels = c("Yes", "No", "Unknown"))

map %>%
  ggplot(aes(x = long, y = lat)) + 
  geom_path(aes(group = group)) + 
  geom_point(data = filter1905T, 
             aes(x = filter1905T$'Longitude location of institution', y = filter1905T$'Latitude location of institution', color = filter1905T$student_writer)) +
  geom_label(data=filter1905, aes(filter1905$'Longitude location of institution', filter1905$'Latitude location of institution', label = year, color = filter1905$student_writer), hjust = 0, nudge_x = 0.3, check_overlap = TRUE) +
  scale_colour_brewer(palette = "Set1") +
  theme(panel.grid = element_blank()) +
  coord_map() +
  theme(legend.position = 'none', axis.text.x = element_blank(),
  axis.text.y = element_blank(), axis.ticks = element_blank()) +
  ylab("") +
  xlab("")


filter <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1915 & join_data2$year > 1910)

filterT <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1915)

filterT$student_writer <- factor(filterT$student_writer, levels = c("Yes", "No", "Unknown"))

map %>%
  ggplot(aes(x = long, y = lat)) + 
  geom_path(aes(group = group)) + 
  geom_point(data = filterT, 
             aes(x = filterT$'Longitude location of institution', y = filterT$'Latitude location of institution', color = filterT$student_writer)) +
  geom_label(data=filter, aes(filter$'Longitude location of institution', filter$'Latitude location of institution', label = year, color = filter$student_writer), hjust = 0, nudge_x = 0.3, check_overlap = TRUE) +
  scale_colour_brewer(palette = "Set1") +
  theme(panel.grid = element_blank()) +
  coord_map() +
  theme(legend.position = 'none', axis.text.x = element_blank(),
  axis.text.y = element_blank(), axis.ticks = element_blank()) +
  ylab("") +
  xlab("")

filter <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1920 & join_data2$year > 1915)

filterT <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1920)

filterT$student_writer <- factor(filterT$student_writer, levels = c("Yes", "No", "Unknown"))

map %>%
  ggplot(aes(x = long, y = lat)) + 
  geom_path(aes(group = group)) + 
  geom_point(data = filterT, 
             aes(x = filterT$'Longitude location of institution', y = filterT$'Latitude location of institution', color = filterT$student_writer)) +
  geom_label(data=filter, aes(filter$'Longitude location of institution', filter$'Latitude location of institution', label = year, color = filter$student_writer), hjust = 0, nudge_x = 0.3, check_overlap = TRUE) +
  scale_colour_brewer(palette = "Set1") +
  theme(panel.grid = element_blank()) +
  coord_map() +
  theme(legend.position = 'none', axis.text.x = element_blank(),
  axis.text.y = element_blank(), axis.ticks = element_blank()) +
  ylab("") +
  xlab("")

filter <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1930 & join_data2$year > 1920)

filterT <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1930)

filterT$student_writer <- factor(filterT$student_writer, levels = c("Yes", "No", "Unknown"))

map %>%
  ggplot(aes(x = long, y = lat)) + 
  geom_path(aes(group = group)) + 
  geom_point(data = filterT, 
             aes(x = filterT$'Longitude location of institution', y = filterT$'Latitude location of institution', color = filterT$student_writer)) +
  geom_label(data=filter, aes(filter$'Longitude location of institution', filter$'Latitude location of institution', label = year, color = filter$student_writer), hjust = 0, nudge_x = 0.3, check_overlap = TRUE)+
  scale_colour_brewer(palette = "Set1") +
  theme(panel.grid = element_blank()) +
  coord_map() +
  theme(legend.position = 'none', axis.text.x = element_blank(),
  axis.text.y = element_blank(), axis.ticks = element_blank()) +
  ylab("") +
  xlab("")

filter <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1940 & join_data2$year > 1930)

filterT <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1940)

filterT$student_writer <- factor(filterT$student_writer, levels = c("Yes", "No", "Unknown"))

map %>%
  ggplot(aes(x = long, y = lat)) + 
  geom_path(aes(group = group)) + 
  geom_point(data = filterT, 
             aes(x = filterT$'Longitude location of institution', y = filterT$'Latitude location of institution', color = filterT$student_writer)) +
  geom_label(data=filter, aes(filter$'Longitude location of institution', filter$'Latitude location of institution', label = year, color = filter$student_writer), hjust = 0, nudge_x = 0.3, check_overlap = TRUE)+
  scale_colour_brewer(palette = "Set1") +
  theme(panel.grid = element_blank()) +
  coord_map() +
  theme(legend.position = 'none', axis.text.x = element_blank(),
  axis.text.y = element_blank(), axis.ticks = element_blank()) +
  ylab("") +
  xlab("")

filter <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1950 & join_data2$year > 1940)

filterT <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1950)

filterT$student_writer <- factor(filterT$student_writer, levels = c("Yes", "No", "Unknown"))

map %>%
  ggplot(aes(x = long, y = lat)) + 
  geom_path(aes(group = group)) + 
  geom_point(data = filterT, 
             aes(x = filterT$'Longitude location of institution', y = filterT$'Latitude location of institution', color = filterT$student_writer)) +
  geom_label(data=filter, aes(filter$'Longitude location of institution', filter$'Latitude location of institution', label = year, color = filter$student_writer), hjust = 0, nudge_x = 0.3, check_overlap = TRUE)+
  scale_colour_brewer(palette = "Set1") +
  theme(panel.grid = element_blank()) +
  coord_map() +
  theme(legend.position = 'none', axis.text.x = element_blank(),
  axis.text.y = element_blank(), axis.ticks = element_blank()) +
  ylab("") +
  xlab("")

filter <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1960 & join_data2$year > 1950)

filterT <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1960)

filterT$student_writer <- factor(filterT$student_writer, levels = c("Yes", "No", "Unknown"))

map %>%
  ggplot(aes(x = long, y = lat)) + 
  geom_path(aes(group = group)) + 
  geom_point(data = filterT, 
             aes(x = filterT$'Longitude location of institution', y = filterT$'Latitude location of institution', color = filterT$student_writer)) +
  geom_label(data=filter, aes(filter$'Longitude location of institution', filter$'Latitude location of institution', label = year, color = filter$student_writer), hjust = 0, nudge_x = 0.3, check_overlap = TRUE)+
  scale_colour_brewer(palette = "Set1") +
  theme(panel.grid = element_blank()) +
  coord_map() +
  theme(legend.position = 'none', axis.text.x = element_blank(),
  axis.text.y = element_blank(), axis.ticks = element_blank()) +
  ylab("") +
  xlab("")

filter <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1970 & join_data2$year > 1960)

filterT <- join_data2 %>%
  filter(join_data2$year, join_data2$year < 1970)

filterT$student_writer <- factor(filterT$student_writer, levels = c("Yes", "No", "Unknown"))

map %>%
  ggplot(aes(x = long, y = lat)) + 
  geom_path(aes(group = group)) + 
  geom_point(data = filterT, 
             aes(x = filterT$'Longitude location of institution', y = filterT$'Latitude location of institution', color = filterT$student_writer)) +
  geom_label(data=filter, aes(filter$'Longitude location of institution', filter$'Latitude location of institution', label = year, color = filter$student_writer), hjust = 0, nudge_x = 0.3, check_overlap = TRUE)+
  scale_colour_brewer(palette = "Set1") +
  theme(panel.grid = element_blank()) +
  coord_map() +
  theme(legend.position = 'none', axis.text.x = element_blank(),
  axis.text.y = element_blank(), axis.ticks = element_blank()) +
  ylab("") +
  xlab("")

map %>%
  ggplot(aes(x = long, y = lat)) + 
  geom_path(aes(group = group)) + 
  geom_point(data = filterT, 
             aes(x = filterT$'Longitude location of institution', y = filterT$'Latitude location of institution', color = filterT$student_writer)) +
  scale_colour_brewer(palette = "Set1") +
  theme(panel.grid = element_blank()) +
  coord_map() +
  theme(legend.position = 'none', axis.text.x = element_blank(),
  axis.text.y = element_blank(), axis.ticks = element_blank()) +
  ylab("") +
  xlab("")
```

```{r}
join_data2 <- join_data2 %>%
  mutate(age = if_else(join_data2$year <= 1922, 'old', 'new'))

join_data2$student_writer <- factor(join_data2$student_writer, levels = c("Yes", "No", "Unknown"))

join_data2 %>%
  group_by(age, student_writer) %>%
  ggplot(aes(age, fill = student_writer), rm.na = TRUE) +
  geom_bar()+
  scale_fill_manual(values=c("red", "blue", "white"))

join_data2 %>%
  group_by(speed_cat) %>%
  ggplot(aes(x = speed_cat, y = year, color = student_writer)) +
  geom_point()+
  geom_boxplot() +
  scale_colour_brewer(palette = "Set1") +
  xlab("Length & Speed")

join_data2 %>%
  ggplot(aes(x = Niche_Athletic_Rank, y = X2019_FB_Wins, color = as.factor(number_fights))) +
  geom_point() +
   scale_colour_brewer(palette = "Set1")

join_data2 %>%
  ggplot(aes(x = join_data2$'Tuition and fees, 2013-14', y = Niche_Athletic_Rank, color = speed_cat)) +
  geom_point()
  
join_data2 %>%
  ggplot(aes(x = Niche_Party_Rank, y = join_data2$'Graduation rate - Bachelor degree within 4 years, total', color = nonsense)) +
  geom_point()

join_data2 %>%
  ggplot(aes(x = men, y = join_data2$'Percent of total enrollment that are women', fill = men)) +
  geom_boxplot() +
  xlab("Does the song refer to either men or boys?") +
  ylab("Percentage of Students who are Female")

join_data2 %>%
  ggplot(aes(x = spelling, y = join_data2$'SAT Writing 75th percentile score', color = spelling)) +
  geom_point(size = 3) +
  scale_colour_brewer(palette = "Set1") +
  xlab("Does the song spell out words?") +
  ylab("SAT Writing 75th Percentile Score")
  

join_data2 %>%
  ggplot(aes(x = spelling, y = join_data2$'SAT Writing 25th percentile score', color = spelling)) +
  geom_point(size = 3) +
  scale_colour_brewer(palette = "Set1") +
  xlab("Does the song spell out words?") +
  ylab("SAT Writing 25th Percentile Score")

plot2 <- join_data2 %>%
  group_by(speed_cat) %>%
  ggplot(aes(x = Niche_Athletic_Rank, y = Niche_Party_Rank, color = speed_cat, text = school)) +
  geom_point()+
  scale_colour_brewer(palette = "Set1") +
  xlab("Athletic Ranking") +
  ylab("Party Ranking")

plotly2 <- ggplotly(plot2, tooltip = "text")

plotly2

join_data_longer <- join_data2 %>%
  pivot_longer(victory:spelling, names_to = 'lyric', values_to='score')

join_data_longer %>%
  group_by(lyric) %>%
  ggplot(aes(x = score, y = X2019_FB_Wins)) +
  geom_boxplot() +
  facet_wrap(~lyric) +
  ylab("2019 Football Wins") +
  xlab("")

join_data_longer %>%
  group_by(lyric) %>%
  ggplot(aes(x = score, fill = score)) +
  geom_bar() +
  facet_wrap(~lyric) +
  xlab("") +
  scale_fill_manual(values=c("red", "green"))
```

----
henrys part
```{r}
join_data %>%
  ggplot(aes(x = sec_duration, y = Niche_Athletic_Rank, color = victory_win_won)) +
  geom_point(aes(size = X2019_FB_Wins)) +
  scale_colour_brewer(palette = "Set2") +
  xlab("Song Length (seconds)") +
  ylab("Athletic Rank")
  
join_data %>%
  ggplot(aes(x = Niche_Athletic_Rank, y = number_fights, color = conference)) +
  geom_point(aes(size = X2019_FB_Wins)) +
  scale_colour_brewer(palette = "Set2") +
  xlab("Athletic Rank") +
  ylab("Number of 'Fights'") +
  ggtitle("Number of 'Fight' Chants vs Athletic Rank")
```

----
matthews part
```{r}
join_data = join_data %>%
  rename(Wins = X2019_FB_Wins, Losses = X2019_FB_Losses)
ggplot(join_data, aes(x = Niche_Athletic_Rank, y = Niche_Party_Rank)) +
  geom_point(aes(size = Wins / 3, shape = student_writer, color = conference)) +
  xlim(-10, 110) +
  ylim(-10, 270) +
  geom_vline(xintercept = 50) +
  geom_hline(yintercept = 120) +
  ggtitle("Athletic rank vs Party rank") +
  xlab("Athletic Rank") +
  ylab("Party Rank") 

ggplot(join_data, aes(x = Losses, y = Wins)) +
  geom_point(aes(color = conference, shape = student_writer, size = join_data$Niche_Party_Rank)) +
  ggtitle("Wins vs Losses")

ggplot(join_data, aes(x = join_data$`ACT Composite 25th percentile score`, y = join_data$Niche_Party_Rank)) +
  geom_point(aes(color = conference, shape = student_writer)) +
  geom_text(label = join_data$school, check_overlap = TRUE) +
  xlab("ACT Composite 25th percentile score") +
  ylab("Party Rank") +
  ggtitle("ACT Composite 25th percentile score vs Party rank")
```

## Party schools fight songs writer varies while party rank < 100 but mostly written by non-students while party rank > 100.
## The biggest Party and athletic schools are from the Big Ten and SEC. The smallest ones are from ACC and the Pac-12.

## Schools with a non student fight song have a better football record than schools with a student writer
## The schools with the best football record also have a low party rank

## Schools between an ACT 25th percentile of 20 and 26 have the highest party rank. Then the party rank increases exponentially as the ACT composite scores exceed 28


----
anns part (if doing more than cleaning)
```{r}

```



